Empirical Bayes for drug efficacy prediction in cell lines

### 2019-04-17 docs/presentation.Rnw
added mcmc version of inverse Gaussian model in stan

>>>>

  \begin{frame}
    \frametitle{Inverse Gaussian distribution}
<<dens_igaussian_igamma, fig.cap="Several densities of the inverse Gaussian and inverse Gamma families", out.width="85%", fig.asp=2/3>>=
digauss <- function(x, theta, lambda, eta) {
  sqrt(lambda/(x^3*2*pi))*exp(-lambda*(x - theta)^2/(2*theta^2*x))
}

digamma <- function(x, theta, lambda, eta) {
  alpha <- eta/2
  beta <- lambda/2
  beta^alpha/gamma(alpha)*x^(-alpha - 1)*exp(-beta/x)
}

eta <- c(1, 2, 1, 1, 5)
lambda <- c(1, 1, 2, 1/2, 2)
theta <- c(1000, 1000, 1, 1, 1)

labels <- c("inv. Gaussian", "inv. Gamma")
col <- c(1, 2)

opar <- par(no.readonly=TRUE)
par(mar=opar$mar*c(1, 1.3, 1, 1), cex=1.2, lwd=2)
layout(matrix(c(rep(c(1, 1, 2, 2, 3, 3), 2), rep(c(0, 4, 4, 5, 5, 0), 2)),
                 nrow=4, ncol=6, byrow=TRUE))
for(i in 1:length(eta)) {
  modes <- c(theta[i]*(sqrt(1 + 9*theta[i]^2/
                            (4*lambda[i]^2)) - 3*theta[i]/(2*lambda[i])),
             lambda[i]/(eta[i] + 2))
  ylim <- c(0, max(digauss(modes[1], theta[i], lambda[i], eta[i]),
                   digamma(modes[2], theta[i], lambda[i], eta[i])))
  curve(digauss(x, theta[i], lambda[i], eta[i]), 0.001, 5, n=1000,
        ylim=ylim, ylab="Density", "x", col=col[1], 
        main=bquote(eta==.(eta[i])*","~lambda==.(lambda[i])*","~theta==.(
          theta[i])))
  curve(digamma(x, theta[i], lambda[i], eta[i]), add=TRUE, 
        col=col[2], n=1000)
}
legend("topright", legend=labels, lty=1, col=col, seg.len=1)
par(opar)
@
  \end{frame}

  \begin{frame}
    \frametitle{Normal inverse Gaussian model}
<<dens_igaussian_marginalbeta, fig.cap="Marginal $\\beta_j$ prior for several choices of $\\gamma^2_d$ hyperpior, scaled to $\\V(\\beta_j)=1$.", out.width="80%", fig.asp=2/3>>=
library(GeneralizedHyperbolic)
# all scaled such that variance is one
sigma <- 1
theta <- 1/sigma^2
lambda <- c(0.1, 2)
lambda1 <- sqrt(2)
col <- c(1:4)
lty <- c(1:4)
labels <- as.expression(c(bquote("IG, "~lambda==.(lambda[1])~", "~
                                   theta==.(theta)),
                          bquote("IG, "~lambda==.(lambda[2])~", "~
                                   theta==.(theta)),
                          "Point mass (ridge)", "Exponential (lasso)"))
dprior <- function(x, lambda, theta, sigma) {
  dnig(x, 0, sigma*sqrt(lambda), sqrt(lambda/(theta*sigma)), 0)
}
dlasso <- function(x, lambda1) {
  0.5*lambda1*exp(-lambda1*abs(x))
}

opar <- par(no.readonly=TRUE)
par(mar=opar$mar*c(1, 1.3, 1/1.3, 1), cex.lab=1.5, lwd=3)
curve(dprior(x, lambda[1], theta, sigma), -3, 3, ylab=expression(p(beta[j])), 
      xlab=expression(beta[j]), n=1000, 
      col=col[1], lty=lty[1])
curve(dprior(x, lambda[2], theta, sigma), add=TRUE, n=1000, 
      col=col[2], lty=lty[2])
curve(dnorm(x, 0, 1), add=TRUE, n=1000, col=col[3], lty=lty[3])
curve(dlasso(x, lambda1), add=TRUE, n=1000, col=col[4], lty=lty[4])
legend("topright", legend=labels, col=col, lty=lty, 
       title="Hyperprior")
par(opar)
@
  \end{frame}
